---
title: "Absorbances Fourth Project"
subtitle: "ContextBase, http://contextbase.github.io"
author: "John Akwei, ECMp ERMp Data Scientist"
date: "January 29, 2016"
output: html_document
---

# Synopsis  
This document is the third analysis of absorbance data involving E.Coli and S.Aureus. Via Data Science analysis, (using the R programming language), the document author intends to build an understanding, numerically and statistically, of the difference between the samples, with p-values, t-tests, etc.  

The dataset examined is: fib_vs_ABS_s_aureus2.xlsx  
4) disc_ecoli.xlsx  
5) disc_SA.xlsx  
6) fill_vs_ecoli.xlsx  
7) fil_vs_S_aureus2.xlsx

The data is re-formatted, in order that each row is a separate record, and the first row contains the variable names.  

The free, open source application RStudio, (version 0.98.1103), was used for the analysis. The version of the R programming language used was, R version 3.2.0 (2015-04-16). The PC operating system is Windows 8.1.  

#### Processing Requirements  
```{r, warning=F, message=F}
setwd("C:/Users/johnakwei/Dropbox/Programming/Morphotypes")
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("xlsx")) {install.packages("xlsx"); require("xlsx")}
```

# Data Input  
```{r, warning=F, message=F}
fib_vs_ABS_s_aureus2 <- read.xlsx("fib_vs_ABS_s_aureus2.xlsx", 1)
names(fib_vs_ABS_s_aureus2) <-c("fib.vs.ABS.S.aureus", "X0mg",
                                "sample.1", "sample.2", "sample.3",
                                "sample.4", "sample.5", "sample.6",
                                "sample.7", "sample.8")
fib.vs.ABS.S.Aureus <- matrix(1,3)
fib.vs.ABS.S.Aureus[1,] <- 1
fib.vs.ABS.S.Aureus[2,] <- 2
fib.vs.ABS.S.Aureus[3,] <- 3
fib_vs_ABS_s_aureus2 <- cbind(fib.vs.ABS.S.Aureus, fib_vs_ABS_s_aureus2)

disc_ecoli <- read.xlsx("disc_ecoli.xlsx", 1)
disc_SA <- read.xlsx("disc_SA.xlsx", 1)
fill_vs_ecoli <- read.xlsx("fill_vs_ecoli.xlsx", 1)
fil_vs_S_aureus2 <- read.xlsx("fil_vs_S_aureus2.xlsx", 1)
```

# Dataset 3 - Graphs, Linear Regression, Correlation, and T-Tests  
```{r, warning=F, message=F}
rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$X0mg, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$X0mg)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Control 0mg") +
  xlab("Samples") + ylab("Control - 0mg - Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~fib_vs_ABS_s_aureus2$X0mg)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$X0mg, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$X0mg)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.1, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.1)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 1") +
  xlab("Samples") + ylab("Sample 1 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.1, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.1, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.1)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.2, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.2)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 2") +
  xlab("Samples") + ylab("Sample 2 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.2, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.2, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.2)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.3, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.3)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 3") +
  xlab("Samples") + ylab("Sample 3 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.3, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.3, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.3)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.4, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.4)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 4") +
  xlab("Samples") + ylab("Sample 4 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.4, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.4, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.4)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.5, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.5)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 5") +
  xlab("Samples") + ylab("Sample 5 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

# fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
#             fib_vs_ABS_s_aureus2$sample.5, na.rm=T)
# layout(matrix(c(1,2,3,4),2,2))
# plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.5, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.5)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.6, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.6)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 6") +
  xlab("Samples") + ylab("Sample 6 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.6, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.6, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.6)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.7, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.7)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 7") +
  xlab("Samples") + ylab("Sample 7 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.7, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.7, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.7)

rLine <- coef(lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
                   fib_vs_ABS_s_aureus2$sample.8, data=fib_vs_ABS_s_aureus2))
layout(matrix(1,1,1))
ggplot(fib_vs_ABS_s_aureus2, aes(x=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
                              y=fib_vs_ABS_s_aureus2$sample.8)) +
  geom_bar(aes(), stat="identity",
           fill=as.factor(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus)) +
  ggtitle("fib_vs_ABS_s_aureus2 Sample 8") +
  xlab("Samples") + ylab("Sample 8 Absorbance") +
  theme(plot.title=element_text(color="blue")) +
  theme(axis.title=element_text(color="darkblue")) +
  geom_abline(intercept=rLine[1], slope=rLine[2], col="#008000")

fit <- lm(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus~
            fib_vs_ABS_s_aureus2$sample.8, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
cor.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
         fib_vs_ABS_s_aureus2$sample.8, method="spearman")
t.test(as.numeric(fib_vs_ABS_s_aureus2$fib.vs.ABS.S.Aureus),
       fib_vs_ABS_s_aureus2$sample.8)
```