---
title: "E.Coli and S.Aureus Absorbances"
author: "John Akwei, ECMp ERMp Data Scientist"
date: "January 19, 2016"
output: html_document
---

# Synopsis  
The objective of this document is to do statistical analysis on absorbance datasets. Via Data Science analysis, (using the R programming language), the document author intends to build an understanding, numerically and statistically, of the difference between the samples, with p-values, t-tests, etc.  

## Table of Contents  
1) Graphs of Data  
2) Linear Regression, Correlation, and T-Tests  

# 1) Graphs of Data  

#### The data is cleaned. Each row is a separate record. The first row contains the variable names. Then the cleaned data is read into R: 

#### Processing Requirements  
```{r, warning=F, message=F}
setwd("C:/Users/johnakwei/Dropbox/Programming/Morphotypes")
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
```

#### Data Input  
```{r, warning=F, message=F}
absorbance1 <- read.csv("Absorbances1.csv", header=T)
absorbance2 <- read.csv("Absorbances2.csv", header=T)
```

#### Exploratory Data Analysis  
```{r, warning=F, message=F}
names(absorbance1)
names(absorbance2)
absorbance1$Powder_mg_vs_Absorbance_E_coli <- as.numeric(absorbance1$Powder_mg_vs_Absorbance_E_coli)
absorbance2$Powder_Mg_vs_S_aureus <- as.numeric(absorbance2$Powder_Mg_vs_S_aureus)
```

## Graphs  
```{r, echo=F, warning=F, message=F}
plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_1,
     main="E.Coli Sample 1 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 1 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_1~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_2,
     main="E.Coli Sample 2 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 2 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_2~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_3,
     main="E.Coli Sample 3 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 3 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_3~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_4,
     main="E.Coli Sample 4 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 4 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_4~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_5,
     main="E.Coli Sample 5 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 5 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_5~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_6,
     main="E.Coli Sample 6 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 6 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_6~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_7,
     main="E.Coli Sample 7 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 7 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance1$Sample_7~absorbance1$Powder_mg_vs_Absorbance_E_coli), col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_1,
     main="S.Aureus Sample 1 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 1 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sample_1~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_2,
     main="S.Aureus Sample 2 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 2 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sample_2~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sampele_3,
     main="S.Aureus Sample 3 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 3 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sampele_3~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_4,
     main="S.Aureus Sample 4 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 4 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sample_4~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_5,
     main="S.Aureus Sample 5 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 5 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sample_5~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_6,
     main="S.Aureus Sample 6 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 6 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sample_6~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

plot(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_7,
     main="S.Aureus Sample 7 Absorbance per MG",
     col.main="blue", xlab="Milligrams (3 measurements per MG)",
     ylab="Sample 7 Absorbance",
     col.lab="darkblue")
abline(lm(absorbance2$Sample_7~absorbance2$Powder_Mg_vs_S_aureus),
       col="#008000")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_1)) +
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 1 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 1") +
labs(fill="Milligrams")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_2)) + 
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 2 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 2") +
labs(fill="Milligrams")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_3)) + 
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 3 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 3") +
labs(fill="Milligrams")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_4)) + 
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 4 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 4") +
labs(fill="Milligrams")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_5)) + 
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 5 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 5") +
labs(fill="Milligrams")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_6)) + 
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 6 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 6") +
labs(fill="Milligrams")

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_7)) + 
geom_boxplot(aes(fill=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli))) +
ggtitle("E.Coli Absorbance per MG - Sample 7 - Medians") +
xlab("Quantities in Milligrams") + ylab("E.Coli Absorbance - Sample 7") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sample_1)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 1 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 1") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sample_2)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 2 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 2") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sampele_3)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 3 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 3") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sample_4)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 4 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 4") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sample_5)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 5 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 5") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sample_6)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 6 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 6") +
labs(fill="Milligrams")

ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                             y=absorbance2$Sample_7)) +
geom_boxplot(aes(fill=as.factor(absorbance2$Powder_Mg_vs_S_aureus))) +
ggtitle("S.Aureus Absorbance per MG - Sample 7 - Medians") +
xlab("Quantities in Milligrams") + ylab("S.Aureus Absorbance - Sample 7") +
labs(fill="Milligrams")
```

#### E.Coli Sample 1 - Sums of 3 Measurements per MG  
```{r, echo=F, warning=F, message=F}
EColi_Sample_1_Sums <- tapply(absorbance1$Sample_1,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_1_Sums

ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_1)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 1') +
ggtitle('E.Coli Sample 1 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

EColi_Sample_2_Sums <- tapply(absorbance1$Sample_2,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_2_Sums
ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_2)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 2') +
ggtitle('E.Coli Sample 2 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

EColi_Sample_3_Sums <- tapply(absorbance1$Sample_3,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_3_Sums
ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_3)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 3') +
ggtitle('E.Coli Sample 3 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

EColi_Sample_4_Sums <- tapply(absorbance1$Sample_4,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_4_Sums
ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_4)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 4') +
ggtitle('E.Coli Sample 4 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

EColi_Sample_5_Sums <- tapply(absorbance1$Sample_5,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_5_Sums
ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_5)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 2') +
ggtitle('E.Coli Sample 5 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

EColi_Sample_6_Sums <- tapply(absorbance1$Sample_6,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_6_Sums
ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_6)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 6') +
ggtitle('E.Coli Sample 6 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

EColi_Sample_7_Sums <- tapply(absorbance1$Sample_7,
absorbance1$Powder_mg_vs_Absorbance_E_coli, FUN=sum, na.rm=T)
EColi_Sample_7_Sums
ggplot(absorbance1, aes(x=as.factor(absorbance1$Powder_mg_vs_Absorbance_E_coli), y=absorbance1$Sample_7)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('E.coli Absorbance Sample 7') +
ggtitle('E.Coli Sample 7 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))
```

#### S.Aureus Sample 1 - Sums of 3 Measurements per MG  
```{r, echo=F, warning=F, message=F}
SAureus_Sample_1_Sums <- tapply(absorbance2$Sample_1,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sample_1_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sample_1)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 1') +
ggtitle('S.Aureus Sample 1 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

SAureus_Sample_2_Sums <- tapply(absorbance2$Sample_2,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sample_2_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sample_2)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 2') +
ggtitle('S.Aureus Sample 2 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

SAureus_Sampele_3_Sums <- tapply(absorbance2$Sampele_3,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sampele_3_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sampele_3)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 3') +
ggtitle('S.Aureus Sample 3 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

SAureus_Sample_4_Sums <- tapply(absorbance2$Sample_4,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sample_4_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sample_4)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 4') +
ggtitle('S.Aureus Sample 4 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

SAureus_Sample_5_Sums <- tapply(absorbance2$Sample_5,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sample_5_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sample_5)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 5') +
ggtitle('S.Aureus Sample 5 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

SAureus_Sample_6_Sums <- tapply(absorbance2$Sample_6,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sample_6_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sample_6)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 6') +
ggtitle('S.Aureus Sample 6 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))

SAureus_Sample_7_Sums <- tapply(absorbance2$Sample_7,
                           absorbance2$Powder_Mg_vs_S_aureus, FUN=sum, na.rm=T)
SAureus_Sample_7_Sums
ggplot(absorbance2, aes(x=as.factor(absorbance2$Powder_Mg_vs_S_aureus),
                        y=absorbance2$Sample_7)) +
geom_bar(aes(), stat="identity") +
xlab('Milligrams') + ylab('S.Aureus Absorbance Sample 7') +
ggtitle('S.Aureus Sample 7 absorbance (Sums of 3 measurements per MG)') +
theme(plot.title=element_text(color="blue")) +
theme(axis.title=element_text(color="darkblue"))
```

# 2) Linear Regression, Correlation, and T-Tests  

#### Linear Regression of E.Coli Sample 1 Data  
```{r, warning=F, message=F}
fit <- lm(absorbance1$Powder_mg_vs_Absorbance_E_coli~
                           absorbance1$Sample_1, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
```

#### Linear Regression of S.Aureus Sample 1 Data  
```{r, warning=F, message=F}
fit <- lm(absorbance2$Powder_Mg_vs_S_aureus~absorbance2$Sample_1, na.rm=T)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
```

#### Correlation of E.Coli Absorbance: Sample 1 - Sample 3:  
```{r, echo=F, warning=F, message=F}
absorbance1_sample1_CorTest <- cor.test(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_1, method="spearman")
absorbance1_sample1_CorTest

absorbance1_sample2_CorTest <- cor.test(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_2, method="spearman")
absorbance1_sample2_CorTest

absorbance1_sample3_CorTest <- cor.test(absorbance1$Powder_mg_vs_Absorbance_E_coli, absorbance1$Sample_3, method="spearman")
absorbance1_sample3_CorTest
```

#### Table of Powder mg vs Absorbance E.coli Correlation P-Values  
```{r, echo=F, warning=F, message=F}
pValueTable <- matrix(nrow=3, ncol=2)
colnames(pValueTable) <- c("Milligrams" ,"P-Value")
rownames(pValueTable) <- c("Sample 1", "Sample 2" , "Sample 3")
pValueTable[1,] <- c("0 Milligrams", absorbance1_sample1_CorTest$p.value)
pValueTable[2,] <- c("1 Milligrams", absorbance1_sample2_CorTest$p.value)
pValueTable[3,] <- c("2 Milligrams", absorbance1_sample3_CorTest$p.value)
pValueTable
```

#### Correlation of S.Aureus Absorbance: Sample 1 - Sample 3:  
```{r, echo=F, warning=F, message=F}
absorbance2_sample1_CorTest <- cor.test(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_1, method="spearman")
absorbance2_sample1_CorTest

absorbance2_sample2_CorTest <- cor.test(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sample_2, method="spearman")
absorbance2_sample2_CorTest

absorbance2_sample3_CorTest <- cor.test(absorbance2$Powder_Mg_vs_S_aureus, absorbance2$Sampele_3, method="spearman")
absorbance2_sample3_CorTest
```

#### Table of Powder mg vs Absorbance S.Aureus Correlation P-Values  
```{r, echo=F, warning=F, message=F}
pValueTable2 <- matrix(nrow=3, ncol=2)
colnames(pValueTable2) <- c("Milligrams" ,"P-Value")
rownames(pValueTable2) <- c("Sample 1", "Sample 2" , "Sample 3")
pValueTable2[1,] <- c("0 Milligrams", absorbance1_sample1_CorTest$p.value)
pValueTable2[2,] <- c("1 Milligrams", absorbance1_sample2_CorTest$p.value)
pValueTable2[3,] <- c("2 Milligrams", absorbance1_sample3_CorTest$p.value)
pValueTable2
```

## T-Tests  
```{r,  echo=F, warning=F, message=F}
E_Coli_Ttest_1 <- t.test(absorbance1$Powder_mg_vs_Absorbance_E_coli,
                         absorbance1$Sample_1)
E_Coli_Ttest_2 <- t.test(absorbance1$Powder_mg_vs_Absorbance_E_coli,
                         absorbance1$Sample_2)
E_Coli_Ttest_3 <- t.test(absorbance1$Powder_mg_vs_Absorbance_E_coli,
                         absorbance1$Sample_3)
S_Aureus_Ttest_1 <- t.test(absorbance2$Powder_Mg_vs_S_aureus,
                           absorbance2$Sample_1)
S_Aureus_Ttest_2 <-t.test(absorbance2$Powder_Mg_vs_S_aureus,
                          absorbance2$Sample_2)
S_Aureus_Ttest_3 <-t.test(absorbance2$Powder_Mg_vs_S_aureus,
                          absorbance2$Sampele_3)
```

#### E.Coli T-Test Comparisons - Table 1  
```{r, echo=F, warning=F, message=F}
EColi_tTestTable <- matrix(nrow=3, ncol=3)
colnames(EColi_tTestTable) <- c("Milligrams" ,"P-Value",
                                "Alternate Hypothesis")
rownames(EColi_tTestTable) <- c("Sample 1", "Sample 2" , "Sample 3")
EColi_tTestTable[1,] <- c("0 Milligrams", E_Coli_Ttest_1$p.value,
                          E_Coli_Ttest_1$alternative)
EColi_tTestTable[2,] <- c("1 Milligrams", E_Coli_Ttest_2$p.value,
                          E_Coli_Ttest_2$alternative)
EColi_tTestTable[3,] <- c("2 Milligrams", E_Coli_Ttest_3$p.value,
                          E_Coli_Ttest_3$alternative)
EColi_tTestTable
```

#### E.Coli T-Test Comparisons - Table 2  
```{r, echo=F, warning=F, message=F}
EColi_tTestTable2 <- matrix(nrow=3, ncol=3)
colnames(EColi_tTestTable2) <- c("Milligrams" , "Confidence Int Low",
                                "Confidence Int High")
rownames(EColi_tTestTable2) <- c("Sample 1", "Sample 2" , "Sample 3")
EColi_tTestTable2[1,] <- c("0 Milligrams", E_Coli_Ttest_1$conf)
EColi_tTestTable2[2,] <- c("1 Milligrams", E_Coli_Ttest_2$conf)
EColi_tTestTable2[3,] <- c("2 Milligrams", E_Coli_Ttest_3$conf)
EColi_tTestTable2
```

#### S.Aureus T-Test Comparisons - Table 1  
```{r, echo=F, warning=F, message=F}
SAureus_tTestTable <- matrix(nrow=3, ncol=3)
colnames(SAureus_tTestTable) <- c("Milligrams" ,"P-Value",
                                "Alternate Hypothesis")
rownames(SAureus_tTestTable) <- c("Sample 1", "Sample 2" , "Sample 3")
SAureus_tTestTable[1,] <- c("0 Milligrams", S_Aureus_Ttest_1$p.value,
                          S_Aureus_Ttest_1$alternative)
SAureus_tTestTable[2,] <- c("1 Milligrams", S_Aureus_Ttest_2$p.value,
                          S_Aureus_Ttest_2$alternative)
SAureus_tTestTable[3,] <- c("2 Milligrams", S_Aureus_Ttest_3$p.value,
                          S_Aureus_Ttest_3$alternative)
SAureus_tTestTable
```

#### S.Aureus T-Test Comparisons - Table 2  
```{r, echo=F, warning=F, message=F}
SAureus_tTestTable2 <- matrix(nrow=3, ncol=3)
colnames(SAureus_tTestTable2) <- c("Milligrams" , "Confidence Int Low",
                                "Confidence Int High")
rownames(SAureus_tTestTable2) <- c("Sample 1", "Sample 2" , "Sample 3")
SAureus_tTestTable2[1,] <- c("0 Milligrams", S_Aureus_Ttest_1$conf)
SAureus_tTestTable2[2,] <- c("1 Milligrams", S_Aureus_Ttest_2$conf)
SAureus_tTestTable2[3,] <- c("2 Milligrams", S_Aureus_Ttest_3$conf)
SAureus_tTestTable2
```
