---
title: "Protein Signal Analysis - Phase 1"
subtitle: "by ContextBase, http://contextbase.github.io"
author: "John Akwei, ECMp ERMp Data Scientist"
date: "February 11, 2016"
output: html_document
---

# Synopsis  
This document is comprised of the following R data analysis items:  
1) Graphs of Protein Data, (with legends), for visualization/understanding the data  
2) Graphs/Tables, (with legends), of the data parameters, (quartiles, etc.)  
3) Hypothesis Tests of Protein Data (correlation coefficients, summary stats, (mean, sd), 95% Confidence)  
4) Notated R Code for future Reproducibility  

The dataset examined is: data_protein.csv  

The following parameters of data_protein.csv are examined:  
1) a - g are the protein signal values (they are not absolute values), but they are relative for each protein separately. They represent the total amount of protein in that sample.  
2) V are the variables measured from each of the 61 samples. v8 - v11 weren't done by me and represent fold changes compared to some control.  
3) V3 is known to be "bad or hi" above 0.7 and is the most important followed by v2 = "bad or hi" at 7 or 8 and above, and then v1 which is considered to be "normal" between 0.86 and 1.76 ug/ml. v3 is the ratio of v2/k (potassium).  
4) Previous tried: Fisher's exact tests by cross tabulation using hi (0.7) v3 group that contained either hi or lo protein vs the rest of the v3 group < 0.7. Samples hi in v3 are more likely than lo to have hi/lo levels of protein c.  

Data Science Goals:  
1) To know if my biomarkers (western blot data/signal intensity) can be used to predict elevated levels of some variable.  
2) To find cutoff points for these markers that could indicate "good" or "bad" (using a certain amount of signal measured).  
3) Find any meaningful correlations with the proteins of interest (POI), n = 7 columns, that could help predict biological dysfunction.  
4) Develop a methodical approach to analyzing the data so I can answer those 2 questions.  

#### Processing Requirements  
```{r, warning=F, message=F}
# Use setwd() to set working directory to folder w/ relevant files
setwd("C:/Users/johnakwei/Dropbox/Programming/Morphotypes")

# Check if required R packages are installed. If not, then install.
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
```

# Data Input  
```{r, warning=F, message=F}
# Read in the .csv datafile from the Working Directory
data_protein <- read.csv("data_protein.csv", header=T, na.strings="NA")
```

# Graphs  
```{r, warning=F, message=F}
# Define Vertical Lines for the first Graph
confLine <- coef(lm(data_protein$a~data_protein$a, data=data_protein))
mLine <- mean(na.omit(data_protein$a))
sdLine <- sd(na.omit(data_protein$a))
sdLines <- c(mLine + sdLine, mLine - sdLine) 
st <- t.test(data_protein$a, data_protein$v3)

# First Graph
ggplot(data_protein,
       aes(x=as.factor(data_protein$a), y=data_protein$a)) + 
geom_point(aes(data_protein$a), col="#00CED1") +
labs(list(title="Protein - Signal 'a' Correlation", x="a",
          y="a", color="darkblue")) +
geom_vline(xintercept=mLine, col="#1E90FF") +
geom_vline(xintercept=sdLines[2], col="#008000") +
geom_vline(xintercept=sdLines[1], col="#008000") +
geom_vline(xintercept=st$conf[1], col="#FF4500") +
geom_vline(xintercept=st$conf[2], col="#FF4500") +
annotate("text", x=c(sdLines[2], sdLines[1], st$conf,
                     mLine), y=c(25, 25, 30, 40, 25),
         label=c("SD Low", "SD High", "95% Conf Low",
                 "95% Conf High", "Mean")) +
theme(plot.title=element_text(vjust=1, color="blue", size=16,
                              face="bold.italic")) +
theme(axis.text.x=element_text(angle=0, hjust=1, color="darkorange",
                               size=14, face="bold")) +
theme(axis.text.y=element_text(color="darkorange", size=14, face="bold")) +
theme(axis.title.x=element_text(color="darkblue", size=14, vjust=0.5),
      axis.title.y=element_text(color="darkblue", size=14, vjust=0.5))
```

```{r, echo=F, warning=F, message=F}
# Create labels for first graph, above.
guideTable <- matrix(nrow=4, ncol=1)
colnames(guideTable) <- c("Signal_Value_a")
rownames(guideTable) <- c("Mean =" ,"SD =",
                          "Conf Low =", "Conf High =")
guideTable[,1] <- c(mLine, sdLine, st$conf[1], st$conf[2])
guideTable <- data.frame(guideTable)
guideTable
```

# Quantiles of 'a' values
```{r, warning=F, message=F}
aQuantile <- quantile(data_protein$a, na.rm=T)
aQuantile
```

# Subset 'a' values
```{r, warning=F, message=F}
# The entire data_protein data is subsetted based on v3 <> 0.7
highValues <- subset(data_protein, data_protein$v3 > 0.7)
highValues <- data.frame(highValues)
lowValues <- subset(data_protein, data_protein$v3 < 0.7)
lowValues <- data.frame(lowValues)
a25quantile <- subset(data_protein, data_protein$a < aQuantile[2])
```

# Plot High 'a' Values
```{r, warning=F, message=F}
# Define Vertical Lines for the High 'a' Values Graph
mLineHigh <- mean(na.omit(highValues$a))
sdLineHigh <- sd(na.omit(highValues$a))
sdLinesHigh <- c(mLineHigh + sdLineHigh, mLineHigh - sdLineHigh) 
stHigh <- t.test(highValues$a, data_protein$v3)

# High 'a' Values Graph
ggplot(highValues,
       aes(x=as.factor(highValues$a), y=highValues$a)) + 
geom_point(aes(highValues$a), col="#00CED1") +
labs(list(title="High 'a' Values Correlation", x="'a' Values > 0.7",
          y="'a' Values > 0.7")) +
geom_vline(xintercept=mLineHigh, col="#1E90FF") +
geom_vline(xintercept=sdLinesHigh[2], col="#008000") +
geom_vline(xintercept=sdLinesHigh[1], col="#008000") +
geom_vline(xintercept=stHigh$conf[1], col="#FF4500") +
geom_vline(xintercept=stHigh$conf[2], col="#FF4500") +
annotate("text", x=c(sdLines[2], sdLines[1], stHigh$conf,
                     mLine), y=c(25, 25, 30, 35, 25),
         label=c("SD Low", "SD High", "95% Conf Low",
                 "95% Conf High", "Mean")) +
theme(plot.title=element_text(vjust=1, color="blue", size=16,
                              face="bold.italic")) +
theme(axis.text.x=element_text(angle=0, hjust=1, color="orange",
                               size=14, face="bold")) +
theme(axis.text.y=element_text(color="orange", size=14, face="bold")) +
theme(axis.title.x=element_text(color="darkblue", size=14, vjust=0.5),
      axis.title.y=element_text(color="darkblue", size=14, vjust=0.5))
```

```{r, echo=F, warning=F, message=F}
# Create labels for high 'a' graph, above.
guideTable2 <- matrix(nrow=4, ncol=1)
colnames(guideTable2) <- c("Signal_Value_a_greater_than_0.7")
rownames(guideTable2) <- c("Mean =" ,"SD =",
                          "Conf Low =", "Conf High =")
guideTable2[,1] <- c(mLineHigh, sdLineHigh, stHigh$conf[1], stHigh$conf[2])
guideTable2 <- data.frame(guideTable2)
guideTable2
```

# Quantiles of High 'a' values
```{r, warning=F, message=F}
aQuantileHigh <- quantile(highValues$a, na.rm=T)
aQuantileHigh
```

# Plot Low 'a' Values
```{r, warning=F, message=F}
# Define Vertical Lines for the Low 'a' Values Graph
mLineLow <- mean(na.omit(lowValues$a))
sdLineLow <- sd(na.omit(lowValues$a))
sdLinesLow <- c(mLineLow + sdLineLow, mLineLow - sdLineLow) 
stLow <- t.test(lowValues$a, data_protein$v3)

# Low 'a' Values Graph
ggplot(lowValues, aes(x=as.factor(lowValues$a), y=lowValues$a)) + 
geom_point(aes(lowValues$a), col="#00CED1") +
labs(list(title="Low 'a' Values Correlation", x="'a' Values < 0.7",
          y="'a' Values < 0.7")) +
geom_vline(xintercept=mLineLow, col="#1E90FF") +
geom_vline(xintercept=sdLinesLow[2], col="#008000") +
geom_vline(xintercept=sdLinesLow[1], col="#008000") +
geom_vline(xintercept=stLow$conf[1], col="#FF4500") +
geom_vline(xintercept=stLow$conf[2], col="#FF4500") +
annotate("text", x=c(sdLines[2], sdLines[1], stLow$conf,
                     mLine), y=c(25, 25, 30, 35, 25),
         label=c("SD Low", "SD High", "95% Conf Low",
                 "95% Conf High", "Mean")) +
theme(plot.title=element_text(vjust=1, color="blue", size=16,
                              face="bold.italic")) +
theme(axis.text.x=element_text(angle=0, hjust=1, color="orange",
                               size=14, face="bold")) +
theme(axis.text.y=element_text(color="orange", size=14, face="bold")) +
theme(axis.title.x=element_text(color="darkblue", size=14, vjust=0.5),
      axis.title.y=element_text(color="darkblue", size=14, vjust=0.5))
```

```{r, echo=F, warning=F, message=F}
# Create labels for low 'a' graph, above.
guideTable3 <- matrix(nrow=4, ncol=1)
colnames(guideTable3) <- c("Signal_Value_a_less_than_0.7")
rownames(guideTable3) <- c("Mean =" ,"SD =",
                          "Conf Low =", "Conf High =")
guideTable3[,1] <- c(mLineLow, sdLineLow, stLow$conf[1], stLow$conf[2])
guideTable3 <- data.frame(guideTable3)
guideTable3
```

```{r, warning=F, message=F, fig.width=11, fig.height=7}
# Graph of a - g Signal Values for all Samples
ggplot(data_protein) +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$a),
             col="Dodger Blue") +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$b),
             col="Forest Green") +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$c),
             col="Gold") +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$d),
             col="Light Coral") +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$e),
             col="Orange Red") +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$f),
             col="Magenta") +
  geom_point(aes(x=as.factor(data_protein$sample), y=data_protein$g),
             col="Medium Orchid") +
  ggtitle("Protein - a through g levels") +
  xlab("Sample") + ylab("a through g Levels") +
  theme(plot.title=element_text(vjust=1, color="blue", size=16,
                                face="bold.italic")) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(axis.text.y=element_text(color="darkorange", size=12, face="bold")) +
  theme(axis.title.x=element_text(color="darkblue", size=14, vjust=0.5),
      axis.title.y=element_text(color="darkblue", size=14, vjust=0.5))
```

# Quantiles of Low 'a' values
```{r, warning=F, message=F}
aQuantileLow <- quantile(lowValues$a, na.rm=T)
aQuantileLow
```

# Linear Regression - Response Variable=Signal "a", Predictor Variable=v3
```{r, warning=F, message=F}
# A linear model is created w/ Signal "a" as the Response variable
# v3 is the Predictor variable
fit <- lm(data_protein$a~data_protein$v3)
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
layout(matrix(1,1,1))
```

# Correlation  
```{r, warning=F, message=F}
# Spearman's test of correlation
ct <- cor.test(data_protein$a, data_protein$v3, method="spearman")
ct
```

# T-Test  
```{r, warning=F, message=F}
# The Welsch Two Sample T-Test
st <- t.test(data_protein$a, data_protein$v3)
st
```