---
title: "ContextBase Predictive Analytics"
author: "Authored by John Akwei, ECMp ERMp Data Scientist"
date: "June 11, 2016"
output: html_document
---

## This document contains examples of the Predictive Analytics capabilities of ContextBase, http://contextbase.github.io.  

## Predictive Analytics Example 1: Linear Regression  

Linear Regression allows for prediction of future occurrences derived from one explanatory variable, and one response variable.  

```{r}
library(knitr)
revenue <- freeny.x
kable(head(revenue))
colnames(revenue) <- c("lag_quarterly_revenue", "price_index",
                       "income_level", "market_potential")
revenue <- data.frame(revenue)
model <- lm(market_potential~price_index, revenue)

cat("The Intercept =", model$coefficients[1])

test <- data.frame(price_index=4.57592)
result <- predict(model, test)

plot(revenue$market_potential, revenue$price_index, col = "red",
     main = "Price Index affecting Market Potential",
     abline(lm(revenue$price_index~revenue$market_potential)),
     cex = 1.3, pch = 16, xlab = "Market Potential",
     ylab = "Price Index")
```

## Example 1 - Linear Regression Conclusion:  
```{r}
cat("For a Price Index of ", as.character(test), ", the predicted Market Potential = ", round(result, 2), ".", sep="")
```

### In conclusion to ContextBase Predictive Analytics Example 1, a direct correlation of Price Index to Market Potential was found, (see above graph). As a test of the Predictive Algorithm, a Price Index of 4.57592 was processed, and a Market Potential of 13.03 was predicted. The source R dataset shows this prediction to be accurate.  

## Predictive Analytics Example 2: Logistic Regression  

Logistic Regression allows for prediction of a logical, (Yes or No), occurrence based on the effects of an explanatory variable on a response variable. For example, the probability of winning a congressional election vs campaign expenditures.

How does the amount of money spent on a campaign affect the probability that the candidate will win the election?  

Source of Data Ranges: https://www.washingtonpost.com/news/the-fix/wp/2014/04/04/think-money-doesnt-matter-in-elections-this-chart-says-youre-wrong/  
```{r}
Expenditures <- c(1000000, 1100000, 1200000, 1300000, 1400000,
                  1500000, 1600000, 1700000, 1800000, 1900000,
                  2000000, 2100000, 2200000, 2300000, 2400000,
                  2000000, 2100000, 2200000, 2300000, 2400000)
ElectionResult <- c(0, 0, 0, 0, 0, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1,	1, 1, 1, 1, 1)

CampaignCosts <- data.frame(Expenditures, ElectionResult)

kable(CampaignCosts, caption = "Campaign Expenditures of Candidates vs Election Results")
```

```{r}
library(ggplot2)
ggplot(CampaignCosts, aes(Expenditures, ElectionResult)) +
  geom_point(aes()) +
  geom_smooth(method='glm',
              method.args = list(family = "binomial"), se=FALSE) +
  labs (x="Campaign Expenditures", y="Probability of Winning Election", title="Probability of Winning Election vs Campaign Costs")
```

The logistic regression analysis gives the following output:  
```{r}
model <- glm(ElectionResult ~., family=binomial(link='logit'), data=CampaignCosts)
model$coefficients
```

The output indicates that campaign expenditures significantly affect the probability of winning the election. The output provides the coefficients for Intercept = -7.615054e+00, and Expenditures = 4.098080e-06. These coefficients are entered in the logistic regression equation to estimate the probability of winning the election:  

Probability of winning election = 1/(1+exp(-(-7.615054e+00+4.098080e-06*CampaignExpenses)))  

For a Candidate that has $1,600,000 in expenditures:  
```{r}
CampaignExpenses <- 1600000
ProbabilityOfWinningElection <- 1/(1+exp(-(-7.615054e+00+4.098080e-06*CampaignExpenses)))

cat("Probability of winning Election = 1/(1+exp(-(-7.615054e+00+4.098080e-06*",
    CampaignExpenses, "))) = ", round(ProbabilityOfWinningElection, 2), ".", sep="")
```

For a Candidate that has $2,100,000 in expenditures:
```{r}
CampaignExpenses <- 2100000
ProbabilityOfWinningElection <- 1/(1+exp(-(-7.615054e+00+4.098080e-06*CampaignExpenses)))

cat("Probability of winning Election = 1/(1+exp(-(-7.615054e+00+4.098080e-06*",
    CampaignExpenses, "))) = ", round(ProbabilityOfWinningElection, 2), ".", sep="")
```

## Example 2 - Logistic Regression Conclusion:  
```{r}
ElectionWinTable <- data.frame(column1=c(1100000, 1400000,
                                         1700000, 1900000,
                                         2300000),
                               column2=
      c(round(1/(1+exp(-(-7.615054e+00+4.098080e-06*1100000))), 2),
      round(1/(1+exp(-(-7.615054e+00+4.098080e-06*1400000))), 2),
      round(1/(1+exp(-(-7.615054e+00+4.098080e-06*1700000))), 2),
      round(1/(1+exp(-(-7.615054e+00+4.098080e-06*1900000))), 2),
      round(1/(1+exp(-(-7.615054e+00+4.098080e-06*2300000))), 2)))
names(ElectionWinTable) <- c("Campaign Expenses", "Probability of Winning Election")
kable(ElectionWinTable)
```

### In conclusion to ContextBase Predictive Analytics Example 2, a direct correlation of Campaign Expenditures to Election Performance was verified. The above table displays corresponding probablities of winning an election to campaign expenses.  

## Predictive Analytics Example 3: Multiple Regression  

Multiple Regression allows for the prediction of the future values of a response variable, based on values of multiple explanatory variables.  

```{r}
input <- data.frame(state.x77[,1:4])
kable(head(input))
colnames(input) <- c("Population", "Income", "Illiteracy", "Life_Exp")

# Create the relationship model.
model <- lm(Life_Exp~Population+Income+Illiteracy, data=input)

# Show the model.
print(model)

# Multiple Regression Plot
par(mfrow=c(1,3))
termplot(model)

a <- coef(model)[1]
cat("The Multiple Regression Intercept = ", a, ".", sep="")

XPopulation <- coef(model)[2]
XIncome <- coef(model)[3]
XIlliteracy <- coef(model)[4]

modelCoef <- data.frame(XPopulation, XIncome, XIlliteracy)
colnames(modelCoef) <- c("Population", "Income", "Illiteracy")
row.names(modelCoef) <- c("Coefficients")
kable(modelCoef)
```

## Multiple Regression Conclusion:   

```{r}
popl <- 3100
Incm <- 5348
Illt <- 1.1

Y = a + popl * XPopulation + Incm * XIncome + Illt * XIlliteracy
cat("For a City where Population = ", popl, ", Income = ", Incm,  ", and Illiteracy = ", Illt, ",
the predicted Life Expectancy is: ", round(Y, 2), ".", sep="")
```

### In conclusion to ContextBase Predictive Analytics Example 3, the multiple variables of "Population", "Income", and "Illiteracy" were used to determine the predicted "Life Expectancy" of an area corresponding to a USA State. For an area with a Population of 3100, a per capita Income Rate of 5348, and an Illiteracy Rate of 1.1, a Life Expectancy of 71.2 years was predicted.  