---
title: "Systematic Vs. Swing Trading ETF Strategies"
date: "May 30, 2017"
output: html_document
---

<br />

## Synopsis  
This document utilizes the QuantMod and PerformanceAnalytics R packages for backtesting of Systematic ETF strategies versus a Swing Trading ETF strategy. The objective is to determine which trading strategy, (in regards to Exchange Traded Funds), has the best Return Rate.  

Backtesting 5 years for 2 Systematic and 1 Swing Trading strategies:  
Backtest the 3 investing strategies, and compare the yearly ROI %.  
Conclusion about the time invested vs passive investment method.  

References:  
http://news.morningstar.com/articlenet/article.aspx?id=777726  
https://quantstrattrader.wordpress.com/2015/05/13/advertising-a-few-systematic-etfs-strictly-of-my-own-volition/  
https://github.com/jironghuang/DCA-and-VA-comparison/blob/master/Stock%20simulations.R  
https://github.com/DavidMSoto/blog/blob/1b98ed0e0543fb0c17749523bf89f460cbb2f570/finances/easysignal_and_BT.R  
https://github.com/bryangoodrich/bryangoodrich.github.io/blob/6e246c23d84460f3ed03d466f2ed4bf7c2cf83a8/_posts/2017-01-01-Daily-Challenges.md  

<br />

```{r, echo=F, eval=F}
setwd("C:/Users/Administrator/Dropbox/Programming/PredictiveAnalytics/TimeSeriesAnalysis/StockTrading/AutomatedTradingStrategies")
```

<br />

## Working Directory, and Required Packages  
```{r, message=F, warning=F}
# First set the Working Directory
# setwd(" ")

# Required Packages
# install.packages("quantmod")
# install.packages("PerformanceAnalytics")
# install.packages("FinCal")
# install.packages("TTR")

library(quantmod)
library(PerformanceAnalytics)
library(FinCal)
library(TTR)
library(ggplot2)
```

<br />

## Strategy 1A:  
### Systematic Trading - Dollar Cost Averaging, (DCA) 
```{r warning=F, message=F}
# Dollar-cost averaging (DCA) is an investment technique of buying a fixed dollar
# amount of a particular investment on a regular schedule, regardless of the
# share price. The investor purchases more shares when prices are low and fewer
# shares when prices are high. The premise is that DCA lowers the average share
# cost over time, increasing the opportunity to profit. The DCA technique does
# not guarantee that an investor won't lose money on investments. Rather, it is
# meant to allow investment over time instead of investment as a lump sum.

# Investing a fixed amount of dollars in a specific security at regular set intervals over a period of time. Dollar cost averaging results in a lower average cost per share, compared with purchasing a constant number of shares at set intervals. The investor buys more shares when the price is low and buys fewer shares when the price is high.

# Select 5 ETFs with Diversified Portfolio Category/Country, and Morningstar rating 4-5 stars:

# Get symbols
symbols <- c("IJH","SCHB","SCHX","VGK","VXF")
getSymbols(symbols, from="2012-01-01")

# Create prices dataframe
prices <- list()
for(i in 1:length(symbols)) {
  prices[[i]] <- Ad(get(symbols[i]))
}
prices <- do.call(cbind, prices)
colnames(prices) <- gsub("\\.[A-z]*", "", colnames(prices))

# Calculate returns
returns <- na.omit(Return.calculate(prices))

# Create a chart of returns
charts.PerformanceSummary(returns, main = "ETF Returns Chart")

# Print Statistics
stats <- rbind(table.AnnualizedReturns(returns),
               maxDrawdown(returns),
               CalmarRatio(returns),
               SortinoRatio(returns) * sqrt(252))
print(round(stats, 3))

# ENTRY STRATEGY:  
# Buy every month into a preselected ETF for $100

# Percent change between start to end
q <- function(start, end) (end-start) / start

x <- Op(to.monthly(prices$IJH))
x <- as.vector(x)
n <- length(x)
final_price <- x[n]
x <- x[-n]

purchased <- ceiling(100 / x)  # monthly shares purchased
invested <- purchased * x  # monthly expense
sum(invested - 100)  # Total amount extra spent above $1,000 each month

sum(invested)  # Total invested (without changes)

sum(purchased)  # Total shares after period

result <- c(
    'dca_result' = sum(q(x, final_price) * invested),
    'lump_result' = sum(invested) * q(x[1], final_price),
    'avg_result' = mean(q(x, final_price)) * sum(invested))

result / sum(invested) * 100  # proportion gained

# EXIT STRATEGY:  
# Each time an ETF/Fund reaches between +15% - 20%
# sell and re-screen for the next best fund
# once it gets above the 15% don't let it get under it, with stop loss
# but if it keeps going up put next stop loss on 20%,
# and keep going every 5%
# Fund/ETF meeting criteria above and start reInvesting every month:

IJHsell <- returns$IJH*1000 > 15 & returns$IJH*1000 < 20
SCHBsell <- returns$SCHB*1000 > 15 & returns$SCHB*1000 < 20
SCHXsell <- returns$SCHX*1000 > 15 & returns$SCHX*1000 < 20
VGKsell <- returns$VGK*1000 > 15 & returns$VGK*1000 < 20
VXFsell <- returns$VXF*1000 > 15 & returns$VXF*1000 < 20

plot(as.numeric(IJHsell), type="n", xlab="Time", ylab="",
     main="ETF Sell Signals")
lines(as.numeric(IJHsell))
lines(as.numeric(returns$IJH*10))
lines(as.numeric(SCHBsell), col="red")
lines(as.numeric(SCHXsell), col="blue")
lines(as.numeric(VGKsell), col="green")
lines(as.numeric(VXFsell), col="yellow")

```

<br />

## Strategy 1B:  
### Systematic Trading - Dollar Cost Averaging, (DCA)  
```{r warning=F, message=F}

# Determine the ETF with the greatest range of 52W high/low:
diffIJH <- diff(range(min(returns$IJH), max(returns$IJH)))
diffSCHB <- diff(range(min(returns$SCHB), max(returns$SCHB)))
diffSCHX <- diff(range(min(returns$SCHX), max(returns$SCHX)))
diffVGK <- diff(range(min(returns$VGK), max(returns$VGK)))
diffVXF <- diff(range(min(returns$VXF), max(returns$VXF)))

FundTrends <- matrix(ncol=1, nrow=5)
rownames(FundTrends) <- c("IJH", "SCHB", "SCHX", "VGK", "VXF")
FundTrends[1,] <- diffIJH
FundTrends[2,] <- diffSCHB
FundTrends[3,] <- diffSCHX
FundTrends[4,] <- diffVGK
FundTrends[5,] <- diffVXF

FundTrends <- FundTrends[order(FundTrends, decreasing = T),]

# ENTRY STRATEGY:
# Buy every month into a preselected ETF/Fund for $100

print("52 week high/low of ETFs")
FundTrends

UpwardFund <- FundTrends[1]
paste("The Fund with the greatest 52 week high/low is", names(UpwardFund))

# EXIT STRATEGY:
# Select the funds/ETFs with the most upward potential through the 52W low/high
# sell and re-screen for the next best fund
# and start reInvesting every month

# Percent change between start to end
q <- function(start, end) (end-start) / start

x <- Op(to.monthly(prices$VGK))
x <- as.vector(x)
n <- length(x)
final_price <- x[n]
x <- x[-n]

purchased <- ceiling(100 / x)  # monthly shares purchased
invested <- purchased * x  # monthly expense
sum(invested - 100)  # Total amount extra spent above $1,000 each month

sum(invested)  # Total invested (without changes)

sum(purchased)  # Total shares after period

result <- c(
    'dca_result' = sum(q(x, final_price) * invested),
    'lump_result' = sum(invested) * q(x[1], final_price),
    'avg_result' = mean(q(x, final_price)) * sum(invested))

result / sum(invested) * 100  # proportion gained
```

<br />

## Conclusion (Strategies 1A and 1B):  
In the tests of DCA with ETF above, the first strategy, (1A), of buying into every month at $100 produce better results than strategy 1B, selecting the fund with the greatest 52 week high/low and then applying DCA. Therefore, incresed upward range of ETFs seems to not have an effect on the DCA strategy.  

The average yearly return of Strategy 1A was 36.8, and the average yearly return of Strategy 1B was 7.0  

<br />

## Strategy 2:  
### Swing Trading with Technical Indicators  
```{r warning=F, message=F}
# Technical indicators (Active), try to catch upward and downward motions by using the invert ETFs:
# Slow Entry / Fast Exit ($VIX Volatility Trackers)
# Time Allocation : -10% in VXX -80% in XIV (inverse) â-10% no action and wait

# $VIX:
# >30 Shorting Downward Trend
# 22 < x < 30 Shorting Downward Trend
# 18 < x < 22 Waiting for long position
# x < 15 and/or x < 18 going Long

# ENTRY STRATEGY:
# Signal line: Crossing of the EMA7 with EMA14, upwards crossing (uptrend), downward (downtrend)
# RSI(5)jumps till > 70 or goes down to < 30 (should always be moving upward)  
# Slow  Stochastic (50,2) : raises through 20-line or start(above the 20-line)
# an upward motion (never pointing down!)
# The MACD(12,26,9)>0: gets close to the Crossing point of the signal line, horizontal high above the zero line or facing upward
# MACD(12,26,9) should be convergent with graph : same trend as ETF graph
# M50 position from the graph of the ETF (above M50 and going upward)
# 
# Select Funds and/or ETF with following criteria:
# Average Volume above >200k

# Our signal is a simple adjusted price over 200 day SMA
signal <- prices > xts(apply(na.omit(prices), 2, SMA, n = 200),
                       order.by=index(na.omit(prices)))
 
# equal weight all assets with price above SMA200
weights <- signal/(rowSums(signal)+1e-16)
 
# With Return.portfolio, need all weights to sum to 1
weights$zeroes <- 1 - rowSums(weights)
returns$zeroes <- 0
 
monthlyWeights <- na.omit(weights[endpoints(weights, on = 'months'),])
weights <- na.omit(weights)
returns <- na.omit(returns)

# EXIT STRATEGY:  
# Signal line: Crossing of the  EMA7 with EMA14, downward crossing (downtrend), downward (downtrend):  
# 1. RSI(5) goes/points down  
# 2. De Slow Stochastic(50,2) starts to point down (if goes under 80-line go check the inverse tracker for entry point!)  
# 3. Descending MACD(12,26,9) goes towards a crossing of the Signal line  
# 4. Starting downward convergence between graph and MACD(12,26,9)  

# Upward trends(Upward-Long)
# MACD(12,26,9)>0 and graph above M50
# RSI(5)>70,
# Slow  Stochastic (50,2) goes up crossing 20-line,  
 
# Downward trends(Inverse-Short):
# MACD(12,26,9)<0 and graph under M50,
# RSI(5)<30,
# Slow  Stochastic (50,2) goes downward crossing 80-line  

macd <- MACD(na.omit(prices$IJH), nFast=12, nSlow=26,
            nSig=9,maType=SMA,percent = FALSE)
signal <- Lag(ifelse(macd$macd < macd$signal, -1, 1))
portfolioReturns <- ROC(na.omit(prices$IJH))*signal

plot(portfolioReturns, main="Portfolio Returns")

table.Drawdowns(portfolioReturns, top=10)

charts.PerformanceSummary(portfolioReturns)

# Print Statistics
portfolioStats <- rbind(table.AnnualizedReturns(portfolioReturns),
               maxDrawdown(portfolioReturns),
               CalmarRatio(portfolioReturns),
               SortinoRatio(portfolioReturns) * sqrt(252))
print(round(portfolioStats, 3))
```

<br />

## Conclusion (Strategy 2)  
The average yearly return % of Strategy 2, during the 5 years, was -0.043. Therefore, the backtests of the 3 strategies seems to determine that DCA is as an effective strategy as use of Technical Indicators for trading ETFs.  