---
title: "Zspread GARCH (1,1) Time Series Analysis"
date: "August 1, 2016"
output: html_document
---

# Synopsis  
This document uses GARCH(1,1) modeling to analyze the Z spread data for FN3.5 MBS bonds for 7/1/2011 through 7/1/2016. The Zspread data is imported, the date format is converted to R format, then a time series is created with zoo package formatting. GARCH (1,1) analysis, and forecasting, is then performed with the rugarch package.  

## Working Directory and System Requirements  
```{r, warning=F, message=F}
setwd("C:/Users/johnakwei/Dropbox/Programming/TimeSeriesAnalysis/IFC")
library(dplyr)
library(ggplot2)
library(stats)
library(zoo)
library(xts)
library(forecast)
library(tseries)

# if (!require("rgl")) { install.packages("rgl"); require("rgl") }
# if (!require("rugarch")) { install.packages("rugarch"); require("rugarch") }

library(rugarch)

# if (!require("betategarch")) { install.packages("betategarch"); require("betategarch") }

library(betategarch)

Zspread <- read.csv("Z spread.csv")

Zspread$Date <- as.Date(Zspread$Date, format="%m/%d/%Y")

plot(Zspread$Date, Zspread$Zspread, type="l", col="orange", main="Plot of Zspread Data")
```

## Time Series of Zspread (Zoo format)  
```{r}
Zspread.z = zoo(x=Zspread, order.by=Zspread$Date)
```

## Zspread.z Formatting Information  
```{r, echo=F}
cat("The data class of Zspread.z is ",class(Zspread.z))

cat("The frequency of the Zspread.z Time Series is ", frequency(Zspread.z))

cat("Zspread.z Time Series: Start = ")
as.Date(start(Zspread.z), origin = "1970-01-01")

cat("Zspread.z Time Series: End = ")
as.Date(end(Zspread.z), origin = "1970-01-01")

cat("First 6 rows of Zspread.z Time Series:")
head(Zspread.z)

cat("Structure of Zspread.z Time Series:")
str(Zspread.z)
```

## Zspread.z Time Series Plots  
```{r}
plot(Zspread$Date, as.factor(Zspread.z$Zspread), type="l",
     col="blue", xlab="Zspread$Date",
     ylab="Zspread.z$Zspread", main="Plot of Zspread.z")
```

## GARCH Modeling  
```{r, warning=F, message=F}
y <- data.frame(as.numeric(Zspread.z[,2]))
rownames(y) <- Zspread.z[,1]

plot(data.frame(Zspread.z[,1], y), type='l',
     xlab="Date", ylab="Zspread.z",
     main="Plot of RUGARCH in-sample volatility estimates")

# create a cluster object
cluster = makePSOCKcluster(15)

spec = ugarchspec()

nrow(expand.grid(GARCH = 1:14, VEX = 0:1, VT = 0:1, Mean = 0:1, ARCHM = 0:2, ARFIMA = 0:1, MEX = 0:1, DISTR = 1:10))

spec = ugarchspec(variance.model = list(model = 'eGARCH', garchOrder = c(1, 1)), distribution = 'std')

fit = ugarchfit(spec, y[1:1000, , drop = FALSE], solver = 'hybrid')

coef(fit)

spec = getspec(fit); spec

plot(fit, which = 1)
plot(fit, which = 2)
plot(fit, which = 3)
plot(fit, which = 4)
plot(fit, which = 7)
plot(fit, which = 8)
plot(fit, which = 9)
plot(fit, which = 10)
plot(fit, which = 12)
```

## 20 days Forecasting - GARCH(1,1) Model (Zoo Time Series Format)   
```{r}
forc1 = ugarchforecast(fit, n.ahead = 20)

U = uncvariance(fit)^0.5; U

plot(forc1, which = 1)
plot(forc1, which = 3)
```

## Backtesting  
```{r}
spec = ugarchspec(variance.model = list(model = 'gjrGARCH', garchOrder = c(1, 1)), distribution = 'jsu')
roll = ugarchroll(spec, y, forecast.length = 20, refit.every = 5, refit.window = 'moving', window.size = 1200, calculate.VaR = FALSE, keep.coef = TRUE, cluster = cluster)
show(roll)
```

```{r, echo=F, eval=F, warning=F, message=F}
# Function that do not run
# setstart(spec) < - list(shape = 5)
# setbounds(spec)

# Below function not found
# setfixed(spec) < - as.list(coef(fit))

# Below "filt2" causes exiting
# filt1 = ugarchfilter(spec, y[1:1200, ], n.old = 1000)
# filt2 = ugarchfilter(spec, y[1001:1200, ])

# Below distribution requires a lot of computer time for processing
# gd = ugarchdistribution(fit, n.sim = 500, recursive = TRUE, recursive.length = 6000, recursive.window = 500, m.sim = 100, solver = 'hybrid', cluster = cluster)
# show(gd)
# plot(gd, which = 1, window = 12)
# plot(gd, which = 2, window = 12)
# plot(gd, which = 3, window = 12)
# plot(gd, which = 4, window = 12)

# Below forecasting methods do not run
# forc2 = ugarchforecast(spec, n.ahead = 20, data = y[1:1000, , drop = FALSE])
# forc3 = ugarchforecast(spec, n.ahead = 1, n.roll = 499, data = sp500ret[1:1500, , drop = FALSE], out.sample = 500)
# f1 = as.data.frame(forc1)
# f2 = as.data.frame(forc2)
# f3 = t(as.data.frame(forc3, which = 'sigma', rollframe = 'all', aligned = FALSE))

# A graphing function not yet operable
# pp.timeplot <-
# function (x, div = "year", type = "l", xlab = "", xnam = NULL, 
#     ...) 
# {
#     if (is.data.frame(x)) 
#         x <- as.matrix(x)
#     if (is.matrix(x)) {
#         matplot(x, axes = FALSE, type = type, xlab = xlab, ...)
#         if (!length(xnam)) {
#             xnam <- row.names(x)
#         }
#         else if (length(xnam) != nrow(x)) {
#             stop("'xnam' is the wrong length")
#         }
#     }
#     else {
#         plot(x, axes = FALSE, type = type, xlab = xlab, ...)
#         if (!length(xnam)) {
#             xnam <- names(x)
#         }
#         else if (length(xnam) != length(x)) {
#             stop("'xnam' is the wrong length")
#         }
#     }
#     axis(2)
#     if (!is.character(div) || length(div) != 1) {
#         stop("'div' must be a single character string -- given has mode ", 
#             mode(div), " and length ", length(div))
#     }
#     div.menu <- c("years", "decades", "months")
#     div.num <- pmatch(div, div.menu, nomatch = 0)
# 	if(div.num == 0) {
# 		stop("unknown or ambiguous value given for 'div' -- possible ",
# 			"values are: ", paste(div.menu, collapse=", "))
# 	}
# 	div <- div.menu[div.num]
#     switch(div, decades = {
#         tim <- substring(xnam, 1, 3)
#         tlab <- paste(unique(tim), "0's", sep = "")
#     }, years = {
#         tim <- substring(xnam, 1, 4)
#         tlab <- unique(tim)
#     }, months = {
# 		tim <- substring(xnam, 1, 7)
# 		tlab <- substring(unique(tim), 6, 7)
# 	})
#     tcount <- cumsum(table(tim))
#     tbrk <- tcount[-length(tcount)] + 0.5
#     axis(1, at = tbrk, labels = rep("", length(tbrk)), tck = 1, 
#         lty = 2, col = 2)
#     axis(1, at = tbrk, labels = rep("", length(tbrk)), lty = 2, 
#         col = 2)
#     maxd <- max(diff(tcount))
#     tcount[length(tcount)] <- tcount[length(tcount) - 1] + maxd
#     axis(1, at = tcount - maxd/2, labels = tlab, tck = 0)
#     box()
# }
```