---
title: "Zspread GARCH (1,1) Time Series Analysis"
date: "August 1, 2016"
output: html_document
---

# Synopsis  
This document uses GARCH(1,1) modeling to analyze the Z spread data for FN3.5 MBS bonds for 7/1/2011 through 7/1/2016. The Zspread data is imported, the date format is converted to R format, then a time series is created with zoo package formatting. GARCH (1,1) analysis, and forecasting, is then performed with the rugarch package.  

## Working Directory and System Requirements  
```{r, warning=F, message=F}
setwd("C:/Users/johnakwei/Dropbox/Programming/TimeSeriesAnalysis/IFC")
library(dplyr)
library(ggplot2)
library(stats)
library(zoo)
library(xts)
library(forecast)
library(tseries)

# if (!require("rgl")) { install.packages("rgl"); require("rgl") }
# if (!require("rugarch")) { install.packages("rugarch"); require("rugarch") }

library(rugarch)

# if (!require("betategarch")) { install.packages("betategarch"); require("betategarch") }

library(betategarch)

Zspread <- read.csv("Z spread.csv")

Zspread$Date <- as.Date(Zspread$Date, format="%m/%d/%Y")

plot(Zspread$Date, Zspread$Zspread, type="l", col="orange", main="Plot of Zspread Data")
```

## Time Series of Zspread (Zoo format)  
```{r}
Zspread.z = zoo(x=Zspread, order.by=Zspread$Date)
```

## Zspread.z Formatting Information  
```{r, echo=F}
cat("The data class of Zspread.z is ",class(Zspread.z))

cat("The frequency of the Zspread.z Time Series is ", frequency(Zspread.z))

cat("Zspread.z Time Series: Start = ")
as.Date(start(Zspread.z), origin = "1970-01-01")

cat("Zspread.z Time Series: End = ")
as.Date(end(Zspread.z), origin = "1970-01-01")

cat("First 6 rows of Zspread.z Time Series:")
head(Zspread.z)

cat("Structure of Zspread.z Time Series:")
str(Zspread.z)
```

## Zspread.z Time Series Plots  
```{r}
plot(Zspread$Date, as.factor(Zspread.z$Zspread), type="l",
     col="blue", xlab="Zspread$Date",
     ylab="Zspread.z$Zspread", main="Plot of Zspread.z")
```

## GARCH Modeling  
```{r, warning=F, message=F}
y <- data.frame(as.numeric(Zspread.z[,2]))
rownames(y) <- Zspread.z[,1]

plot(data.frame(Zspread.z[,1], y), type='l',
     xlab="Date", ylab="Zspread.z",
     main="Plot of RUGARCH in-sample volatility estimates")

# create a cluster object
cluster = makePSOCKcluster(15)

spec = ugarchspec()

nrow(expand.grid(GARCH = 1:14, VEX = 0:1, VT = 0:1, Mean = 0:1, ARCHM = 0:2, ARFIMA = 0:1, MEX = 0:1, DISTR = 1:10))

spec = ugarchspec(variance.model = list(model = 'eGARCH', garchOrder = c(1, 1)), distribution = 'std')

fit = ugarchfit(spec, y[1:1000, , drop = FALSE], solver = 'hybrid')

coef(fit)

spec = getspec(fit); spec

plot(fit, which = 1)
plot(fit, which = 2)
plot(fit, which = 3)
plot(fit, which = 4)
plot(fit, which = 7)
plot(fit, which = 8)
plot(fit, which = 9)
plot(fit, which = 10)
plot(fit, which = 12)
```

## 20 days Forecasting - GARCH(1,1) Model (Zoo Time Series Format)   
```{r}
forc1 = ugarchforecast(fit, n.ahead = 20)

U = uncvariance(fit)^0.5; U

plot(forc1, which = 1)
plot(forc1, which = 3)
```

## Backtesting  
```{r}
spec = ugarchspec(variance.model = list(model = 'gjrGARCH', garchOrder = c(1, 1)), distribution = 'jsu')

roll = ugarchroll(spec, y, forecast.length = 20, refit.every = 5, refit.window = 'moving', window.size = 1200, calculate.VaR = FALSE, keep.coef = TRUE, cluster = cluster)

show(roll)
```