---
title: "Example of Logistic Regression"
author: "John Akwei, ECMp ERMp Data Scientist"
date: "April 02, 2016"
output: html_document
---

Most of the following text is from the Wikipedia page on Logistic Regression, (https://en.wikipedia.org/wiki/Logistic_regression). The R code in this document was authored by John Akwei, Data Scientist, in order to demonstrate R programming for the Logistic Regression results, and graphs.

#### Logistic Regression  

Example: Probability of passing an exam versus hours of study  

A group of 20 students spend between 0 and 6 hours studying for an exam. How does the number of hours spent studying affect the probability that the student will pass the exam?  
```{r}
Hours <- c(0.50, 0.75, 1.00, 1.25, 1.50, 1.75, 1.75, 2.00, 2.25,
           2.50, 2.75, 3.00, 3.25, 3.50,	4.00,	4.25,	4.50,	4.75,
           5.00, 5.50)
Pass	<- c(0, 0, 0, 0, 0, 0, 1,	0, 1, 0, 1,	0, 1, 0, 1,	1, 1,	1, 1, 1)

HrsStudying <- data.frame(Hours, Pass)
```

The table shows the number of hours each student spent studying, and whether they passed (1) or failed (0).   

```{r}
HrsStudying_Table <- t(HrsStudying); HrsStudying_Table
```
The graph shows the probability of passing the exam versus the number of hours studying, with the logistic regression curve fitted to the data.  

```{r}
library(ggplot2)
ggplot(HrsStudying, aes(Hours, Pass)) +
  geom_point(aes()) +
  geom_smooth(method='glm', family="binomial", se=FALSE) +
  labs (x="Hours Studying", y="Probability of Passing Exam",
        title="Probability of Passing Exam vs Hours Studying")
```

Graph of a logistic regression curve showing probability of passing an exam versus hours studying  

The logistic regression analysis gives the following output.  
```{r}
model <- glm(Pass ~.,family=binomial(link='logit'),data=HrsStudying)
model$coefficients
```

Coefficient	Std.Error	z-value	P-value (Wald)  
Intercept	-4.0777	1.7610	-2.316	0.0206  
Hours	1.5046	0.6287	2.393	0.0167  

The output indicates that hours studying is significantly associated with the probability of passing the exam (p=0.0167, Wald test). The output also provides the coefficients for Intercept = -4.0777 and Hours = 1.5046. These coefficients are entered in the logistic regression equation to estimate the probability of passing the exam:  

Probability of passing exam =1/(1+exp(-(-4.0777+1.5046* Hours)))  

For example, for a student who studies 2 hours, entering the value Hours =2 in the equation gives the estimated probability of passing the exam of p=0.26:  
```{r}
StudentHours <- 2
ProbabilityOfPassingExam <- 1/(1+exp(-(-4.0777+1.5046*StudentHours)))
ProbabilityOfPassingExam
```

Probability of passing exam =1/(1+exp(-(-4.0777+1.5046*2))) = 0.26.  

Similarly, for a student who studies 4 hours, the estimated probability of passing the exam is p=0.87:  
Probability of passing exam =1/(1+exp(-(-4.0777+1.5046*4))) = 0.87.  
```{r}
StudentHours <- 4
ProbabilityOfPassingExam <- 1/(1+exp(-(-4.0777+1.5046*StudentHours)))
ProbabilityOfPassingExam
```

This table shows the probability of passing the exam for several values of hours studying.  
```{r}
ExamPassTable <- data.frame(column1=c(1, 2, 3, 4, 5),
                            column2=c(1/(1+exp(-(-4.0777+1.5046*1))),
                                      1/(1+exp(-(-4.0777+1.5046*2))),
                                      1/(1+exp(-(-4.0777+1.5046*3))),
                                      1/(1+exp(-(-4.0777+1.5046*4))),
                                      1/(1+exp(-(-4.0777+1.5046*5)))))
names(ExamPassTable) <- c("Hours of study", "Probability of passing exam")
ExamPassTable
```