---
title: "Backtesting of 5 year USA Swap Rate Trading Strategies"
date: "August 23, 2016"
output: html_document
---

## Synopsis  
This document uses Systematic Investor Toolbox modeling to backtest the USA Swap Rate data from 7/20/2011 through 7/20/2016.  

The strategies to backtest are;  

SMA (Simple Moving Average) strategy  
Buy Rule: If price exceeds 10 day SMA, buy  
Exit Rule: if price falls below 10 day SMA, sell.  

RSI (Relative Strength Indicator)  
Buy Rule: if 5 day RSI falls below 20, buy  
Exit Rule: If 5 day RSI exceeds 80, sell  
Stop loss order = 5 Take profit order = 10  
Transaction cost: variable, starting from 0.5 bp  

```{r, echo=F}
setwd("C:/Users/johnakwei/Dropbox/Programming/IFC")
```

## Working Directory, and Required Packages  
```{r, message=F, warning=F}
# Set Working Directory
# setwd(" ")

# Required Packages
# if (!require("quantmod")) { install.packages("quantmod"); require("quantmod") }

# if (!require("PerformanceAnalytics")) { install.packages("PerformanceAnalytics"); require("PerformanceAnalytics") }

library(RCurl)
library(zoo)
library(xts)
library(quantmod) # data, plotting, quant modelling
library(PerformanceAnalytics) # performance and risk management
library(knitr)
```

## Load Systematic Investor Toolkit  
```{r, warning=F, message=F}
sit <- getURLContent('https://github.com/systematicinvestor/SIT/raw/master/sit.gz', binary = TRUE, followlocation = TRUE, ssl.verifypeer = FALSE)
con <- gzcon(rawConnection(sit, 'rb'))
source(con)
close(con)
```

## Systematic Investor Analysis    
```{r, warning=F, message=F}
# Create a new environment
data <- new.env()

# Specify the name of the asset and where the csv file is located on your computer.
tickers <- spl('Swap')
file.path <- 'Original data.csv'

Swap_data <- read.csv(file.path, header=T) 
swap_date <- as.Date(Swap_data$Date, format="%d-%b-%y")
swap_date <- format(swap_date, "20%y-%m-%d")

swap_data_zoo <- zoo(as.numeric(paste(
  Swap_data$X10.Years.Mid.Rate.in.bps)), order.by=swap_date)
index(swap_data_zoo) <- as.POSIXct(index(swap_data_zoo))
swap_data_ready <- xts(coredata(swap_data_zoo), order.by = index(swap_data_zoo))

for(n in tickers) {data[[n]] = swap_data_ready}

bt.prep(data, align='remove.na')

# Specify the prices and store our models
prices <- data$Swap
data$prices <- prices
data$execution.price <- prices
models <- list()

# Create Buy and Hold Strategy
data$weight[] = NA
data$weight[] = 1

models$buy.hold <- bt.run.share(data, clean.signal = T,
                               trade.summary = F)

# Calculate the indicators we need for our strategy
CCI20 <- CCI(na.omit(prices), 20) 
RSI3 <- RSI(na.omit(prices), 3)
DEMA10 <- DEMA(na.omit(prices), n = 10, v = 1, wilder = FALSE)
DEMA10c <- prices - DEMA10  
DEMA10c <- DEMA10c/.0001  

#Set our long entry conditions found by our algorithms and optimized by us in the last post
# buy.signal <- ifelse(RSI3 < 30 & CCI20 > -290 & CCI20 < -100 & DEMA10c > -40 & DEMA10c < -20, 1, NA)

buy.signal <- ifelse(RSI3 < 30 & CCI20 > -290 & CCI20 < -100 & DEMA10c > -40 & DEMA10c < 750, 1, NA)

#need to define capital in bt.run()
data$weight[] = NA
data$weight[1:length(buy.signal)] = buy.signal

# Create our long model
models$long <- bt.run.share(data, clean.signal = T,
                            trade.summary = F)

# Set our short conditions
# sell.signal <- ifelse(DEMA10c > 10 & DEMA10c < 40 & CCI20 > 185 & CCI20 < 325 & RSI3 > 50, -1 ,NA)

sell.signal <- ifelse(DEMA10c > 5 & DEMA10c < 200 & CCI20 > 100 & CCI20 < 500 & RSI3 > 10, -1 ,NA)

data$weight[] = NA
data$weight[1:length(sell.signal)] = sell.signal

# Create our short model
models$short <- bt.run.share(data, clean.signal = T,
                             trade.summary = F)  ##share strategy

# Set the long and short conditions for our strategy
# long.short.strategy <- iif(RSI3 < 30 & CCI20 > -290 & CCI20 < -100 & DEMA10c > -40 & DEMA10c < -20,1,iif(DEMA10c > 10 & DEMA10c < 40 & CCI20 > 185 & CCI20 < 325 & RSI3 > 50, -1 ,NA))

long.short.strategy <- iif(RSI3 < 100 & CCI20 > -400 & CCI20 < -50 & DEMA10c > -100 & DEMA10c < 20,1,iif(DEMA10c > 5 & DEMA10c < 200 &
CCI20 > 100 & CCI20 < 1000 & RSI3 > 10, -1 ,NA))

data$weight[] = NA
data$weight[1:length(long.short.strategy)] = long.short.strategy

# Create our long short strategy
models$longshort <- bt.run.share(data, clean.signal = T,
                                 trade.summary = F)

# Isolate the dates from our validation set (The data not used to train the model or create the rules, our out-of-sample test)
dates = '2014-02-26::2014-09-22'
# dates = '2011-07-01::2016-07-01'

# View a plot of our trades
bt.stop.strategy.plot(data, models$longshort, dates = dates, layout=T, main = 'Long Short Strategy', plotX = F)

# View the equity curve and performance statistics.
strategy.performance.snapshoot(models, T)

# The stop loss function
stop.loss <- function(weight, price, tstart, tend, pstop) {
  index = tstart : tend
  if(weight > 0)
    price[ index ] < (1 - pstop) * price[ tstart ]
  else
    price[ index ] > (1 + pstop) * price[ tstart ]
}

# Set our maximum loss at a .25% move in price against our trade
Stoploss = .25/100
# Stoploss = 5

data$weight[] = NA

csf <- custom.stop.fn(coredata(long.short.strategy), coredata(prices), stop.loss,pstop = Stoploss)

data$weight[1:length(csf)] = csf

# Our long short model with a .25% stop loss
models$stoploss <- bt.run.share(data, clean.signal = T,
                                trade.summary = F)

# The plot of our trades
bt.stop.strategy.plot(data, models$stoploss, dates = dates,
                      layout = T, main = 'Stop Loss', plotX = F)

# And how it compares to the original model
strategy.performance.snapshoot(models[c(1,4:5)], T) 

# The take profit function
take.profit <- function(weight, price, tstart, tend, pprofit) {
  index = tstart : tend
  if(weight > 0)
    price[ index ] > (1 + pprofit) * price[ tstart ]
  else
    price[ index ] < (1 - pprofit) * price[ tstart ]
}

# Maintain at 1:1 risk/reward ratio and set our take profit at a .25% change in price
Takeprofit = .25/100
# Takeprofit = 10

data$weight[] = NA

csf_lss <- custom.stop.fn(coredata(long.short.strategy), coredata(prices), take.profit, pprofit = Takeprofit)

data$weight[1:length(csf_lss)] = csf_lss

# Our long short model with a .25% take profit
models$takeprofit <- bt.run.share(data, clean.signal = T,
                                  trade.summary = F)

# The plot of our trades
bt.stop.strategy.plot(data, models$takeprofit, dates = dates, layout=T, main = 'Take Profit', plotX = F)

# Compare it to our other models
strategy.performance.snapshoot(models[c(1,4:6)], T) 

# The stop loss and take profit function
stop.loss.take.profit<-function(weight, price, tstart, tend, pstop, pprofit) {
  index = tstart : tend
  if(weight > 0) {
    temp = price[ index ] < (1 - pstop) * price[ tstart ]
    
    # profit target
    temp = temp | price[ index ] > (1 + pprofit) * price[ tstart ]
  } else {
    temp = price[ index ] > (1 + pstop) * price[ tstart ]
    
    # profit target
    temp = temp | price[ index ] < (1 - pprofit) * price[ tstart ]
  }
  return( temp )
}

data$weight[] = NA

csf_st <- custom.stop.fn(coredata(long.short.strategy), coredata(prices), stop.loss.take.profit,pstop = Stoploss, pprofit = Takeprofit)

data$weight[1:length(csf_st)] = csf_st

# Our long short model with a .25% stop loss and .25% take profit
models$stop.loss.take.profit = bt.run.share(data, clean.signal = T, trade.summary = F)

# The plot of our trades
layout(1:4)
bt.stop.strategy.plot(data, models$longshort, dates = dates, layout=T, main = 'Long Short', plotX = F)
bt.stop.strategy.plot(data, models$stoploss, dates = dates, layout=T, main = 'Long Short .25% SL', plotX = F)
bt.stop.strategy.plot(data, models$takeprofit, dates = dates, layout=T, main = 'Long Short .25% TP', plotX = F)
bt.stop.strategy.plot(data, models$stop.loss.take.profit, dates = dates, layout=T, main = 'Long Short .25% SL, .25% TP', plotX = F)

# Finally comparing all the models we created
strategy.performance.snapshoot(models[c(1,4:7)], T)
```