---
title: "Dashboard"
author: "John Akwei, ECMp ERMp Data Scientist"
date: "Tuesday, November 03, 2015"
output: html_document
---

# Proprietary Business Dashboard example by ContextBase
```{r fig.height=4,echo=FALSE,results='asis'}
cat('Last updated on:',as.character(Sys.time()))
```

```{r fig.height=4, fig.width=12, echo=FALSE}
grid.arrange(plot_sales,plot_profits,nrow=1)
```

Some additional text to add context to the KPI interpretation.
The .R file:

https://rstudio.github.io/shinydashboard/
library(knitr)
library(ggplot2)
library(gridExtra)

## load the data

```{r}
# library(xlsx)
# dataset<-read.xlsx('dataset.xlsx',1)
```

dataset <- matrix(nrow=20, ncol=4)

colnames(dataset) <- c("Year", "Sales", "Profits", "Costs")

dataset[3,] <- c(2002, 490, 85.46, 404.5)
dataset[4,] <- c(2003, 425, 76.84, 348.2)
dataset[5,] <- c(2004, 480, 89.86, 390.1)
dataset[6,] <- c(2005, 480, 82.18, 397.8)
dataset[7,] <- c(2006, 405, 71.93, 333.1)
dataset[8,] <- c(2007, 430, 66.74, 363.3)
dataset[9,] <- c(2008, 485, 83.03, 402.0)
dataset[10,] <- c(2009, 475, 88.16, 386.8)
dataset[11,] <- c(2010, 425,  55.76, 369.2)
dataset[12,] <- c(2011, 430,  69.49, 360.5)

dataset <- data.frame(dataset)

## visualize the data

plot_sales <- ggplot(dataset,aes(x=factor(Year),y=Sales,fill=Sales)) +
xlab('Year') + ylab('Sales ($)') +
geom_bar(stat='identity',colour='black') +
scale_fill_gradient(low='red',high='green') +
ggtitle('Yearly sales') +
theme(axis.text.x = element_text(angle = -35))
  
plot_profits <- ggplot(dataset,aes(x=factor(Year),y=Profits,fill=Profits))+
xlab('Year') + ylab('Profits ($)') +
geom_bar(stat='identity',colour='black') +
scale_fill_gradient(low='red',high='green') +
ggtitle('Yearly profits') +
theme(axis.text.x = element_text(angle = -35))

knit2html('dashboard.Rmd') # produces the HTML file