---
title: "Propensity Score Matching of Insurance Data"
author: "John Akwei, Data Science Specialist"
date: "Friday, September 04, 2015"
output: html_document
---

## Synopsis  
The document performs Propensity Score Matching, in order to find balance between treatment and control groups within a sample of insurance data.  
  
#### Required R language packages for Matched Sampling:  
```{r, message=F, warning=F}
# install.packages("MatchIt")
# install.packages("Matching")
library(MatchIt)
library(Matching)
library(data.table)
require(MatchIt)
require(Matching)
```

#### Import of Insurance Data, and data summary:  
```{r}
dt <- read.csv("Aim1.csv")
dt[1:5,]
```

#### Insurance Data Covariates:  
```{r}
names(dt)
```

#### Subset the Covariates for Matched Sampling:  
```{r}
attach(dt)
dt2 <- cbind(dt[1], dt[13], dt[15:22], dt[25:30], dt[35:36])
names(dt2)
```

#### Propensity Score Matching of Insurance Data, (Treatment & Control Groups):  
```{r}
m.out <- matchit(insurance ~ ID + routine_checkup + great_health + mother_college + HS_grad + low_income + disability + female + NE_So_region + high_exercise + smoker + race_white + daily_veg + daily_fruit + low_risk_alcohol + overweight_11 + obese_11, data=dt2, method="subclass")
# , ratio=1
summary(m.out)
```

#### Create output data table:  
```{r}
matchedIns <- match.data(m.out)
write.csv(matchedIns, file="MatchedData.csv")
```

#### Jitter Plot of Propensity Score Matching:  
```{r, message=F, warning=F}
plot(m.out, type="jitter")
```

#### Histogram of Propensity Score Matching:  
```{r, message=F, warning=F}
plot(m.out, type="hist")
```

## T-tests of Predictor Group Balance  
```{r, echo=F, warning=F, message=F}
t.test(insurance~great_health, matchedIns)
t.test(insurance~mother_college, matchedIns)
t.test(insurance~HS_grad, matchedIns)
t.test(insurance~low_income, matchedIns)
```